%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I, the copyright holder of this work, release this work into the
%% public domain. This applies worldwide. In some countries this may
%% not be legally possible; if so: I grant anyone the right to use
%% this work for any purpose, without any conditions, unless such
%% conditions are required by law.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[
  digital, %% This option enables the default options for the
           %% digital version of a document. Replace with `printed`
           %% to enable the default options for the printed version
           %% of a document.
  notable,   %% Causes the coloring of tables. Replace with `notable`
           %% to restore plain tables.
  nolof,     %% Prints the List of Figures. Replace with `nolof` to
           %% hide the List of Figures.
  nolot,     %% Prints the List of Tables. Replace with `nolot` to
           %% hide the List of Tables.
  nocover
  %% More options are listed in the user guide at
  %% <http://mirrors.ctan.org/macros/latex/contrib/fithesis/guide/mu/fi.pdf>.
]{fithesis3}
%% The following section sets up the locales used in the thesis.
\usepackage[resetfonts]{cmap} %% We need to load the T2A font encoding
\usepackage[T1,T2A]{fontenc}  %% to use the Cyrillic fonts with Russian texts.
\usepackage[
  main=english, %% By using `czech` or `slovak` as the main locale
                %% instead of `english`, you can typeset the thesis
                %% in either Czech or Slovak, respectively.
  english,czech %% The additional keys allow
]{babel}        %% foreign texts to be typeset as follows:
%%
%%   \begin{otherlanguage}{german}  ... \end{otherlanguage}
%%   \begin{otherlanguage}{russian} ... \end{otherlanguage}
%%   \begin{otherlanguage}{czech}   ... \end{otherlanguage}
%%   \begin{otherlanguage}{slovak}  ... \end{otherlanguage}
%%
%% For non-Latin scripts, it may be necessary to load additional
%% fonts:
\usepackage{paratype}
\def\textrussian#1{{\usefont{T2A}{PTSerif-TLF}{m}{rm}#1}}
\newcommand{\symbiotic}{\textsc{Symbiotic}\xspace}
%%
%% The following section sets up the metadata of the thesis.
\thesissetup{
    date          = \the\year/\the\month/\the\day,
    university    = mu,
    faculty       = fi,
    type          = mgr,
    author        = Martina Vitovská,
    gender        = f,
    advisor       = {doc. RNDr. Jan Strejček, Ph.D.},
    title         = {Instrumentation of LLVM IR},
    TeXtitle      = {Instrumentation of LLVM IR},
    keywords      = {instrumentation, static analysis, program analysis, LLVM, memory safety},
    TeXkeywords   = {instrumentation, static analysis, program analysis, LLVM, memory safety},
    thanks        = {Most of all, I would like to thank my advisor Jan Strejček and my consultant Marek Chalupa for their guidance, patience and the time they spent discussing the thesis with me. I would also like to thank my family and friends for their support when I was working on this thesis.},
    abstract      = {The thesis is focused on instrumentation of LLVM IR. First, we give an insight into existing tools that instrument a code to perform an analysis on it. Then we implement and describe a new tool for instrumentation of LLVM IR that is configurable by a~user and can use information from various static analyses. We create a configuration for checking memory safety and reduce the number of inserted checks by employing a pointer analysis and dividing it into phases. At the end, we evaluate the instrumentation for checking memory safety integrated in a verification tool called \symbiotic on a set of benchmarks. },
}

\usepackage{caption}
\newcommand{\thenalg}{\thechapter .\arabic{nalg}}
\DeclareCaptionLabelFormat{algocaption}{Algorithm \thenalg} % defines a new caption label as Algorithm x.y

\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}

\usepackage{makeidx}      %% The `makeidx` package contains
\makeindex                %% helper commands for index typesetting.
%% These additional packages are used within the document:
\usepackage{paralist} %% Compact list environments
\usepackage{amsmath}  %% Mathematics
\usepackage{amsthm}
\usepackage[table]{xcolor}
\usepackage{amsfonts}
\usepackage{url}      %% Hyperlinks
\usepackage{listings} %% Source code highlighting
\lstset{
  basicstyle      = \ttfamily,%
  identifierstyle = \color{black},%
  keywordstyle    = \color{blue},%
  keywordstyle    = {[2]\color{cyan}},%
  keywordstyle    = {[3]\color{olive}},%
  stringstyle     = \color{teal},%
  commentstyle    = \itshape\color{magenta}}
\usepackage{floatrow} %% Putting captions above tables
\floatsetup[table]{capposition=top}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usepackage{styles/llvm}
\usepackage{styles/nasm}

\usepackage{xspace}
\usepackage{xcolor}
\usepackage{subcaption}

\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\llvm}{\textsc{llvm}\xspace}
\newcommand{\klee}{\textsc{Klee}\xspace}
\newcommand{\stacklist}{\texttt{stack\_list}\xspace}
\newcommand{\globalslist}{\texttt{globals\_list}\xspace}
\newcommand{\heaplist}{\texttt{heap\_list}\xspace}
\newcommand{\dealloclist}{\texttt{deallocated\_list}\xspace}
\newcommand{\llvmpin}{\textsc{LLVMPin}\xspace}
\newcommand{\clang}{\textsc{Clang}\xspace}

\lstnewenvironment{algorithm}[1][] %defines the algorithm listing environment
{
    \refstepcounter{nalg} %increments algorithm number
    \captionsetup{labelformat=algocaption,labelsep=colon} %defines the caption setup for: it ises label format as the declared caption label above and makes label and caption text to be separated by a ':'
    \lstset{ %this is the stype
        mathescape=true,
        frame=lines,
        numbers=left,
        numberstyle=\tiny,
		numbersep=5pt,
        basicstyle=\footnotesize,
        keywordstyle=\color{black}\bfseries,
        keywords={,input, output, return, datatype, in, if, else, foreach, while, begin, end, then,and, or, }
        numbers=left,
        xleftmargin=1em,
		framexleftmargin=1em,
        #1 % this is to add specific settings to an usage of this environment (for instnce, the caption and referable label)
    }
}
{}

\begin{document}
\chapter{Introduction}
\input{chapters/01_intro}

\chapter{LLVM}\label{chap:llvm}
\input{chapters/02_llvm.tex}

\chapter{Overview of Existing Tools}\label{chap:tools}
\input{chapters/03_existing_tools.tex}

\chapter{Configurable Instrumentation}\label{chap:instr}
\input{chapters/04_instr.tex}

\chapter{Memory Safety Instrumentation}\label{chap:memsafety}
\input{chapters/05_memsafety.tex}

\chapter{Evaluation}\label{chap:eval}
\input{chapters/06_evaluation.tex}

\chapter{Future Work}\label{chap:future}
\input{chapters/07_future_work.tex}

\chapter{Conclusion}\label{chap:conclusion}
\input{chapters/08_conclusion.tex}

\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography

\appendix %% Start the appendices.
\chapter{An appendix}
Here you can insert the appendices of your thesis.


\end{document}

