As our lives became highly dependent on computers, the question of software
reliability turned to be a serious issue. If we take into consideration the
size and the complexity of today's computer programs, it is almost impossible
to find all the bugs in a code manually. The need for an automatic program
analysis is therefore increasing.

We can distiguish two types of program analysis: dynamic and static. Whereas
dynamic analysis is performed by executing a program, static analysis
inspects a representation of a program (e.g.~a~source code) and analyzes it
without actually running it.

One of the techniques used in the context of both dynamic and static program
analysis is instrumentation. Instrumenting a program means inserting a new code
into an existing code in order to gather information relevant for a program
analysis. The new code usually does not change the behaviour of the program.
Instrumentation is used for example in profilers which insert instructions for
gathering information about time or memory consumption. It is also used by many
tools that check memory safety of programs as it is possible to track the state
of the memory with the inserted code.

The aim of this thesis is to present an overview of existing tools for
instrumentation of LLVM IR bitcode and to design a new tool for
instrumentation of LLVM IR. The tool should be configurable and should offer the
possibility to use results of external static analyses to reduce the amount of
newly inserted code. Another goal is to create a~configuration for finding
memory safety errors such as double free, invalid dereference, or
use-after-free-error.

The new tool is implemented as an independent module, which means it can be
combined with any tool that performs the actual analysis on the instrumented
program. We employed the instrumentation in \symbiotic~\cite{Symbiotic}, which
is an open-source tool for static analysis of sequential programs in LLVM.
\symbiotic is based on three techniques: instrumentation, program slicing, and
symbolic execution. Instrumentation is used to reduce the problem of checking
some property violations (e.g. checking for NULL pointer dereference) to
a~reachability problem. In other words, a given program is instrumented such
that the original program violates the considered property if and only if some
error location is reachable in the instrumented program. Program
slicing~\cite{weiser} is a technique that removes the code that is not relevant
for the reachability of error locations. Therefore, a subsequent analysis is
performed on a smaller amount of code and is thus faster. In \symbiotic,
the sliced program is passed to a symbolic executor~\cite{King} that goes
through every possible execution path of a program and looks whether some
error location is reachable.

\section{Structure of the thesis}

The structure of this thesis is following: In Chapter~\ref{chap:llvm} we give
an insight into LLVM and introduce basic instructions of LLVM IR.
Chapter~\ref{chap:tools} lists the existing tools for instrumenting LLVM IR and
tools that use instrumentation of LLVM IR for analysis of programs. We discuss
benefits and drawbacks of these tools. In Chapter~\ref{chap:instr} the general
approach of our configurable instrumentation is described together with the
configuration files that must be provided. Chapter~\ref{chap:memsafety} deals
with a~configuration of instrumentation for checking memory safety errors.
Besides the basic approach, we also propose an enhancement with a~pointer
analysis and with staged instrumentation to make the whole verification process
faster. Chapter~\ref{chap:eval} presents experimental evaluation of the
instrumentation designed for memory safety implemented in \symbiotic. We
compare the original and the enhanced versions of instrumentation on a set of
benchmarks from SV-COMP
2018~\footnote{\url{https://sv-comp.sosy-lab.org/2018/benchmarks.php}}.  In
Chapter~\ref{chap:future} we suggest future development of the implemented tool
and configurations. Chapter~\ref{chap:conclusion} is a conclusion of this
thesis.
